Supporting Information for Coherent IR-Hyper-Raman Four Wave Mixing Spectroscopy

### 2024-09-22 si.tex
general solution

\section{Evaluating Herzberg-Teller Integrals for Harmonic Wells of Identical Curvature}
In this section, we discuss the evaluation of Herzberg-Teller integrals when the ground and first excited state are harmonic wells of identical curvature.\cite{HerzbergTeller1933}
% We note that Myers et al. have developed a general expression to evaluate Franck-Condon factors for two displaced harmonic wells of differing curvature,\cite{Myers1982} 
We investigate the one dimensional case using the Hamiltonian described in the main text, $H = H_g |g) \left(g| + H_e |e\right) (e|$, where
\begin{subequations}\label{Hamiltonian}
	\begin{equation}
		H_g = \frac{\hbar \omega }{2} \left(p^2 + q^2 \right)
	\end{equation}
	and
	\begin{equation}
		H_e = \frac{\hbar \omega }{2} \left(p^2 +  (q-\Delta)^2 \right) + \hbar \omega_{eg}.
	\end{equation} 
\end{subequations}
Here, $p,q$ are the dimensionless conjugate normal mode coordinates, $\hbar\omega_{eg}$ is the energy difference between $H_e$ and $H_g$, and $\Delta$ is the dimensionless offset of the excited state surface relative to the ground state.
The Hamiltonians are the well-known quantum harmonic oscillator, and their solutions are
\begin{equation}
	\psi_m(q) = \langle q | m \rangle = N_m H_m(q) \exp(-q^2/2),
\end{equation}
where $N_m = (m! \ 2^m \pi^{\frac{1}{2}})^{-\frac{1}{2}}$ is the normalization factor and $H_m$ is the $m^{th}$ Hermite Polynomial.\cite{RN230}

@@

The matrix elements of \autoref{key} can be computed numerically, but analytic forms of the integral also exist. 
Due to the polynomial nature of $H_v(q)$, evaluation of \autoref{key} reduces into evaluating sums of integrals of the form
\begin{equation}\label{integral}
	\int_{-\infty}^\infty \mathrm{d}q \ q^w \exp(-q^2 + q\Delta)
\end{equation}
When $w=0$, the result follows by completing the square:\footnote{$\int_{-\infty}^{\infty} \mathrm{d}x \exp(-(x-a)^2) = \sqrt{\pi}$}
\begin{equation}
	\begin{split}
		\int_{-\infty}^\infty \mathrm{d}q \exp(-q^2 + q\Delta) &= \int_{-\infty}^\infty \mathrm{d}q \exp(-(q^2 - q\Delta +\frac{\Delta^2}{4} - \frac{\Delta^2}{4}))\\
		&= \exp(-\frac{\Delta^2}{4}) \int_{-\infty}^\infty \mathrm{d}q \exp(-(q - \frac{\Delta}{2})^2) \\
		&= \sqrt{\pi} \exp(-\frac{\Delta^2}{4})
	\end{split}
\end{equation}
 For $w \neq 0$, \autoref{integral} is evaluated through repeated use of a well-known Gaussian identity
\begin{equation}\label{identity}
	\int_{-\infty}^\infty \mathrm{d}q \ q^w \exp(-q^2 + q\Delta) = \partial_{\Delta} \int_{-\infty}^\infty \mathrm{d}q \ q^{w-1} \exp(-q^2 + q\Delta)
\end{equation}

We now have the background needed to calculate $\mel{\tilde{1}}{q}{0}$ for an arbitrary offset $\Delta$.
Using $H_0(q) = 1$ and $H_1(q) = 2q$,
\begin{equation}\label{HT10}
	\begin{split}
		\mel{\tilde{1}}{q}{0} &= N_1 N_0 \exp(-\frac{\Delta^2}{2}) \int_{-\infty}^\infty \mathrm{d}q \ 2(q - \Delta) q \exp(-q^2 + q\Delta)\\
		&= 2N_1 N_0 \exp(-\frac{\Delta^2}{2}) \left( \int_{-\infty}^\infty \mathrm{d}q \ q^2 \exp(-q^2 + q\Delta) + \Delta \int_{-\infty}^\infty \mathrm{d}q \ q \exp(-q^2 + q\Delta) \right)
	\end{split}
\end{equation}
Both integrals can be evaluated using \autoref{identity}:
\begin{subequations}\label{evalGaussians}
	\begin{equation}
		\begin{split}
					\int_{-\infty}^\infty \mathrm{d}q \ q \exp(-q^2 + q\Delta) &= \partial_{\Delta} \left(\sqrt{\pi} \exp(\frac{\Delta^2}{4})\right)\\
					& = \sqrt{\pi} \frac{\Delta}{2} \exp(\frac{\Delta^2}{4})
		\end{split}
	\end{equation}
	\begin{equation}
	\begin{split}
		\int_{-\infty}^\infty \mathrm{d}q \ q^2 \exp(-q^2 + q\Delta) &= \partial_{\Delta} \left(\sqrt{\pi} \left(\frac{\Delta}{2} \exp(\frac{\Delta^2}{4})\right)\right)\\
		& =\sqrt{\pi} \left(\frac{1}{2}\exp(\frac{\Delta^2}{4}) + \frac{\Delta^2}{4} \exp(\frac{\Delta^2}{4})\right)
	\end{split}
\end{equation}
\end{subequations}
Substituting the results of \autoref{evalGaussians} into \autoref{HT10}, 
\begin{equation}
	\mel{\tilde{1}}{q}{0} = \sqrt{\frac{1}{2}} (1 - \frac{\Delta^2}{2}) \exp(\frac{-\Delta^2}{4})
\end{equation}
The evaluation of other Herzberg-Teller integrals is analogous.

\pagebreak
\section{Impact of displacement, resonance frequencies, and vibronic linewidth and on HDFG Spectra}
\begin{figure}[!htbp]
	\centering
	\includegraphics[width=6.675in]{figures/fcht.png}
	\caption{Plots of select (a) Franck-Condon factors and (b) Herzberg-Teller integrals relevant to this manuscript as a function of offset, $\Delta$.} 
	\label{fig:fcht}
\end{figure}

>>>>

\section{Evaluating Herzberg-Teller Integrals for Harmonic Wells of Identical Curvature}

In this section, we discuss the evaluation of Herzberg-Teller integrals when the ground and first excited state are harmonic wells of identical curvature.\cite{HerzbergTeller1933}
% We note that Myers et al. have developed a general expression to evaluate Franck-Condon factors for two displaced harmonic wells of differing curvature,\cite{Myers1982} 
We investigate the one dimensional case using the Hamiltonian described in the main text, $H = H_g |g) \left(g| + H_e |e\right) (e|$, where
\begin{subequations}\label{Hamiltonian}
	\begin{equation}
		H_g = \frac{\hbar \omega }{2} \left(p^2 + q^2 \right)
	\end{equation}
	and
	\begin{equation}
		H_e = \frac{\hbar \omega }{2} \left(p^2 +  (q-\Delta)^2 \right) + \hbar \omega_{eg}.
	\end{equation} 
\end{subequations}
Here, $p,q$ are the dimensionless conjugate normal mode coordinates, $\hbar\omega_{eg}$ is the energy difference between $H_e$ and $H_g$, and $\Delta$ is the dimensionless offset of the excited state surface relative to the ground state.
The Hamiltonians are the well-known quantum harmonic oscillator, and their solutions are
\begin{equation}
	\psi_m(q) = \langle q | m \rangle = N_m H_m(q) \exp(-q^2/2),
\end{equation}
where $N_m = (m! \ 2^m \pi^{\frac{1}{2}})^{-\frac{1}{2}}$ is the normalization factor and $H_m$ is the $m^{th}$ Hermite Polynomial.\cite{RN230}

@@

\subsection{General Solution to $\langle \tilde{n} | q^k | m \rangle$}
The matrix elements of \autoref{key} can be computed numerically, but analytic forms of the integral also exist. 
Due to the polynomial nature of $H_v(q)$, the product $H_n(q-\Delta) q^k H_m(q)$ will also be a polynomial, so that the integral of equation \autoref{key} will be of the form
\begin{eqnarray}
	\int dq \ P_{nmk}(q) \exp(-q^2 + q\Delta), \\
	P_{nmk}(q) \equiv \sum_\lambda^{n+m+k} a_\lambda q^\lambda
\end{eqnarray}
for some set coefficients $a_\lambda$, to be specified later. We can recast the integral into the linear combination of integrals,
\begin{equation}\label{eq:analytic_form}
	\mel{\tilde{n}}{q^k}{m} = N_n N_m \exp(-\frac{\Delta^2}{2}) \sum_\lambda a_\lambda F_\lambda(\Delta),
\end{equation}
where
\begin{equation}\label{integral}
	F_w(\Delta) \equiv \int_{-\infty}^\infty \mathrm{d}q \ q^w \exp(-q^2 + q\Delta).
\end{equation}
\autoref{integral} can be readily evaluated, as we now show.  When $w=0$, the integral is evaluated by completing the square:\footnote{$\int_{-\infty}^{\infty} \mathrm{d}x \exp(-(x-a)^2) = \sqrt{\pi}$}
\begin{equation}
	\begin{split}
		F_0(\Delta) &= \int_{-\infty}^\infty \mathrm{d}q \exp(-(q^2 - q\Delta +\frac{\Delta^2}{4} - \frac{\Delta^2}{4}))\\
		&= \exp(\frac{\Delta^2}{4}) \int_{-\infty}^\infty \mathrm{d}q \exp(-(q - \frac{\Delta}{2})^2) \\
		&= \sqrt{\pi} \exp(\frac{\Delta^2}{4})
	\end{split}
\end{equation}
For the general solution, \autoref{integral} is evaluated via repeated use of a well-known Gaussian identity
\begin{equation}\label{identity}
	\begin{split}
		F_w(\Delta) &= \int_{-\infty}^\infty \mathrm{d}q \ q^w \exp(-q^2 + q\Delta) = \partial_{\Delta} \int_{-\infty}^\infty \mathrm{d}q \ q^{w-1} \exp(-q^2 + q\Delta) \\
		&= \partial_\Delta F_{w-1}(\Delta)
	\end{split}
\end{equation}
so we can write
\begin{equation}\label{eq:F_solved}
	\begin{split}
		F_w(\Delta) &= \left(\partial_\Delta \right)^w F_0(\Delta) \\
		&= \sqrt{\pi} \frac{\mathrm{d}^w}{\mathrm{d}\Delta^w} \exp(\frac{\Delta^2}{4}) \\
		&= \sqrt{\pi} \left( -\frac{i}{2} \right)^w \exp(\frac{\Delta^2}{4}) H_w\left(i\frac{\Delta}{2}\right).
	\end{split}
\end{equation}
In the last line we have used the generator definition of Hermite polynomials:  $H_n(x) \equiv (-1)^n e^{x^2} \frac{\mathrm{d}^n}{\mathrm{d}x^n} e^{-x^2}$.

We now will evaluate the polynomial coefficients $P(q)$.
The Hermite polynomials of \autoref{key} can be explicitly written as 
\begin{equation}
\begin{split}
	H_m(q) &= \sum_{x=0}^{\left\lfloor{m/2}\right\rfloor} a_m^{(m-2x)} q^{m-2x} \\
	H_n(q-\Delta) &= \sum_{x=0}^{\left\lfloor{n/2}\right\rfloor}
	a_n^{(n-2x)} (q-\Delta)^{n-2x} \\
		&= \sum_{x=0}^{\left\lfloor{n/2}\right\rfloor} a_n^{(n-2x)} \sum_{l=0}^{n-2x} \Delta^{n-2x-l} \binom{n-2x}{l} q^l
\end{split}
\end{equation}
where $\left\lfloor n \right\rfloor$ denotes the \textit{floor} of $n$, $\binom{a}{b} = a! / (b!(a-b)!)$, and the Hermite coefficients are
\begin{equation}
	a_n^{(n-2x)} = n! \frac{(-1)^x 2^{n-2x}}{x!(n-2x)!}.
\end{equation}
Multiplying these terms together give us explicit polynomial coefficients:
\begin{equation}\label{eq:poly_solved}
\begin{split}
	P_{nmk}(q) %&= \sum_{x=0}^{\left\lfloor{m/2}\right\rfloor} a_m^{(m-2x)} \sum_{y=0}^{\left\lfloor{n/2}\right\rfloor} a_n^{(n-2y)} \sum_{l=0}^{n-2y} \Delta^{n-2y-l} \binom{n-2y}{l} q^{l+m-2x+k} \\
	&= \sum_{x=0}^{\left\lfloor{m/2}\right\rfloor} \sum_{y=0}^{\left\lfloor{n/2}\right\rfloor} \sum_{l=0}^{n-2y} c_{xyl} \ q^{l+m-2x+k}, \\
	c_xyl &= a_m^{(m-2x)} a_n^{(n-2y)} \Delta^{n-2y-l} \binom{n-2y}{l}.
\end{split}
\end{equation}
\autoref{eq:poly_solved} is not in standard polynomial form, because our summation is not singular and coefficients in our summation may have the same power. 
Further simplifications can be made to sum to reduce recounting indices, but we do not find these simplifications essential.
Regardless, substituting \autoref{eq:F_solved} and \autoref{eq:poly_solved} into \autoref{eq:analytic_form}, we can now evaluate the solution as:
\begin{equation}\label{eq:general_solution}
	\mel{\tilde{n}}{q^k}{m} = \sqrt{\pi} N_n N_m \exp(-\frac{\Delta^2}{4}) \sum_{x=0}^{\left\lfloor{m/2}\right\rfloor} \sum_{y=0}^{\left\lfloor{n/2}\right\rfloor} \sum_{l=0}^{n-2y} \left( -\frac{i}{2} \right)^{l+m-2x+k} c_{xyl} H_{l+m-2x+k}\left(i\frac{\Delta}{2}\right).
\end{equation}

\subsection{Evaluating Herzberg Teller terms for HDFG}

\begin{equation}
\begin{split}
	\langle n |q| 0 \rangle &= \cdots \\
\end{split}
\end{equation}

\begin{equation}
	\begin{split}
		\langle n |q| 1 \rangle &= \cdots \\
	\end{split}
\end{equation}

% We now have the background needed to calculate $\langle \tilde{n} |q| m \rangle$.
% % $\mel{\tilde{1}}{q}{0}$ for an arbitrary offset $\Delta$.
% Using $H_0(q) = 1$ and $H_1(q) = 2q$,
% \begin{equation}\label{HT10}
% 	\begin{split}
% 		\mel{\tilde{1}}{q}{0} &= N_1 N_0 \exp(-\frac{\Delta^2}{2}) \int_{-\infty}^\infty \mathrm{d}q \ 2(q - \Delta) q \exp(-q^2 + q\Delta)\\
% 		&= 2N_1 N_0 \exp(-\frac{\Delta^2}{2}) \left( \int_{-\infty}^\infty \mathrm{d}q \ q^2 \exp(-q^2 + q\Delta) + \Delta \int_{-\infty}^\infty \mathrm{d}q \ q \exp(-q^2 + q\Delta) \right)
% 	\end{split}
% \end{equation}
% Both integrals can be evaluated using \autoref{identity}:
% \begin{subequations}\label{evalGaussians}
% 	\begin{equation}
% 		\begin{split}
% 					\int_{-\infty}^\infty \mathrm{d}q \ q \exp(-q^2 + q\Delta) &= \partial_{\Delta} \left(\sqrt{\pi} \exp(\frac{\Delta^2}{4})\right)\\
% 					& = \sqrt{\pi} \frac{\Delta}{2} \exp(\frac{\Delta^2}{4})
% 		\end{split}
% 	\end{equation}
% 	\begin{equation}
% 	\begin{split}
% 		\int_{-\infty}^\infty \mathrm{d}q \ q^2 \exp(-q^2 + q\Delta) &= \partial_{\Delta} \left(\sqrt{\pi} \left(\frac{\Delta}{2} \exp(\frac{\Delta^2}{4})\right)\right)\\
% 		& =\sqrt{\pi} \left(\frac{1}{2}\exp(\frac{\Delta^2}{4}) + \frac{\Delta^2}{4} \exp(\frac{\Delta^2}{4})\right)
% 	\end{split}
% \end{equation}
% \end{subequations}
% Substituting the results of \autoref{evalGaussians} into \autoref{HT10}, 
% \begin{equation}
% 	\mel{\tilde{1}}{q}{0} = \sqrt{\frac{1}{2}} (1 - \frac{\Delta^2}{2}) \exp(\frac{-\Delta^2}{4})
% \end{equation}
% The evaluation of other Herzberg-Teller integrals is analogous.

\newpage
\section{Impact of displacement, resonance frequencies, and vibronic linewidth and on HDFG Spectra}
\begin{figure}[!htbp]
	\centering
	\includegraphics[width=6.675in]{figures/fcht.png}
	\caption{Plots of select (a) Franck-Condon factors and (b) Herzberg-Teller integrals relevant to this manuscript as a function of offset, $\Delta$.} 
	\label{fig:fcht}
\end{figure}
